Tests of measurement invariance without subgroups: A~generalization of classical methods

### 2012-06-08 www/paper.Rnw
Various small touch-ups and improvements. Should be ready to go now.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@134 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

To employ the LR~test for assessing measurement invariance, model
parameters are estimated separately for a certain number of subgroups
of the data (with some parameters potentially restricted to be equal across
subgroups).  For ease of exposition, we describe
the case where there are no such parameter restrictions; as shown in
the example and simulation below, however, it is
straightforward to extend all methods to the more general case.
After fitting the model to each subgroup, the sum of maximized likelihoods 
from the subgroups are compared with the original maximized full-sample likelihood
in a $\chi^2$~test. For the special case of two subgroups,
the alternative $H_1^*$ from (\ref{eq:h1*}) with fixed and prespecified $\nu$ is adopted
and the null hypothesis $H_0$ from (\ref{eq:h0}) reduces to ${\bm \theta}^{(A)} = {\bm \theta}^{(B)}$.
The parameter estimates $\hat {\bm \theta}^{(A)}$ can then be obtained from the
observations $i = 1, \dots, m$, say, for which $v_i \le \nu$. Analogously,
$\hat {\bm \theta}^{(B)}$ is obtained by maximizing the likelihood for the
observations $i = m + 1, \dots, n$, for which $v_i > \nu$. The LR~test statistic
for the given threshold $\nu$ is then
\begin{equation} \label{eq:lr}
  \mathit{LR}(\nu) ~=~ -2 \left[
         \ell(\hat {\bm \theta}; x_1, \dots, x_n)
   ~-~ \{\ell(\hat {\bm \theta}^{(A)}; x_1, \dots, x_m)
    +    \ell(\hat {\bm \theta}^{(B)}; x_{m+1}, \dots, x_n)\}
    \right],
\end{equation}
which, when \eqref{eq:h1*} holds, has an asymptotic $\chi^2$ with degrees
of freedom equal to the number of parameters in~${\bm \theta}$.

@@

For application to measurement invariance, the most important
theoretical result involves the fact that, under the $H_0$ in \eqref{eq:h0}, the
\emph{cumulative score process} converges to a specific asymptotic
process. The $k$-dimensional cumulative score process is defined as
\begin{equation} \label{eq:cumscore}
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n t \rfloor} {\bm s}(\hat {\bm \theta}; x_i)
  \qquad (0 \le t \le 1)
\end{equation}
where $\lfloor nt \rfloor$ is the integer part of $nt$ and
$\hat I$ is some consistent estimate of the covariance matrix of the scores,
e.g., their outer product or the observed information matrix.
The $k$~dimensions of the 
process arise from the fact that a separate cumulative score is
maintained for each of the $k$ model parameters.  
As the equation shows, the cumulative score process adds subsets of
casewise score contributions across individuals along the ordering
w.r.t.\ the variable $V$ of interest.  At $t=1/n$, only the
first individual's contribution enters into the summation; at
$t=2/n$, the first two individuals' contributions enter into the
summation, etc., until $t = n/n$
where all contributions enter into the summation. Thus, due to (\ref{eq:ml1}),
the cumulative score process always equals zero at $t = 0$ and
returns to zero at $t = 1$. Furthermore, multiplication by
$\hat {\bm I}^{-1/2}$ ``decorrelates'' the $k$~cumulative score
processes, such that each univariate process (i.e., each
process for a single model parameter) is unrelated to (and asymptotically independent of) all other
processes. Therefore, this cumulative process ${\bm B}(t; \hat {\bm \theta})$
accomplishes the challenges discussed at the beginning of this section:
it makes use of the ordering of the observations by taking cumulative sums,
and it decorrelates the contributions of the $k$ different parameters.

@@

The empirical cumulative score process from (\ref{eq:cumscore}) can also be viewed
as an $n \times k$ matrix with elements ${\bm B}(i/n; \hat {\bm \theta})_j$ that
we also denote ${\bm B}(\hat {\bm \theta})_{ij}$ below for brevity.
That is, assuming that individuals are ordered by $V$, the first row
of ${\bm B}(\hat {\bm \theta})$ corresponds to the first individual's
scores.  The second row of ${\bm B}(\hat {\bm \theta})$ corresponds to
the sum of the first two individuals' scores, etc., until the last row
of ${\bm B}(\hat {\bm \theta})$ corresponds to the sum of all
individuals' scores (which will be a row of zeroes).
Under this setup, each column of ${\bm B}(\hat {\bm \theta})$
converges to a univariate Brownian bridge and pertains 
to a single factor analysis parameter. To carry out a test of $H_0$, the process/matrix
needs to be aggregated to a scalar test statistic by collapsing across
rows (individuals) and columns (parameters) of the matrix. The asymptotic distribution
of this test statistic is then easily obtained by applying the same
aggregation to the asymptotic Brownian bridge \cite{HjoKon02,ZeiHor07},
so that corresponding $p$-values can be derived.

As argued above, no single aggregation function will have high power for all
conceivable patterns of measurement invariance ${\bm \theta}(V)$, while any
(reasonable) aggregation function will have non-trivial power under $H_1$.
Thus, various aggregation strategies should be employed depending on which
pattern ${\bm \theta}(V)$ is most plausible (because the exact pattern
is typically
unknown). A particularly agnostic aggregation strategy is to reject $H_0$
if any component of the the cumulative score process ${\bm B}(t; \hat {\bm \theta})$
strays ``too far'' from zero at any time, i.e., if
\begin{equation}
    \label{eq:dmax}
    \mathit{DM} = \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,
\end{equation}
becomes large.  In determining where in ${\bm B}(\hat {\bm \theta})$
this maximum occurred, we are able to determine  
threshold(s) of parameter change (over the individuals $i = 1, \dots, n$)
and the parameter(s) affected by it (over $j = 1, \dots, k$). This test is
especially useful for visualization, as the cumulative score process for
each individual parameter can be displayed along with the appropriate
critical value. For an example of this see Figure~\ref{fig:lambdas}
(in the ``Example'' section) which shows the the cumulative score processes for three
factor loadings along with the critical values at 5\% level.

@@

As pointed out above, specification of the asymptotic distribution under $H_0$ for the test
statistics from the previous section is straightforward: it is simply the aggregation
of the asympotic process ${\bm B}^0(t)$ \cite{HjoKon02,ZeiHor07}. Thus, $\mathit{DM}$
from (\ref{eq:dmax}) converges to $\sup_t || {\bm B}^0(t) ||_{\infty}$, where
$||\cdot||_\infty$ denotes the maximum norm. Similarly, $\mathit{CvM}$
from (\ref{eq:cvm}) converges to $\int_0^1 || {\bm B}^0(t) ||_2^2 d t$, where $||\cdot||^2$
denotes the Euclidean norm. Finally, $\max \mathit{LM}$ from (\ref{eq:maxlm}) -- and analogously
the maximum Wald and LR~tests -- converges to $\sup_t (t (1-t))^{-1} || {\bm B}^0(t) ||_2^2$
(which can also be interpreted as the maximum of a tied-down Bessel process, as pointed out
previously).

@@

\subsubsection{Overall Results}
Test statistics for the hypotheses \eqref{eq:simh0} and~\eqref{eq:h0}
are displayed in Figure~\ref{fig:cusumex}. Each panel displays a test 
statistic's fluctuation across values of student age, with the first
column containing tests of~\eqref{eq:simh0} and the second column
containing tests of~\eqref{eq:h0}.  The solid
horizontal 
lines represent critical values for $\alpha= 0.05$, and
the Cram\'{e}r-von Mises panels 
also contain a dashed line depicting the value of the test
statistic (test statistics for the others are simply 
the maxima of the processes).
In other words, for panels in the first and third rows, \eqref{eq:simh0} is
rejected if the process crosses the horizontal line.  For panels in
the second row, \eqref{eq:h0} is rejected if the dashed horizontal line is
higher than the solid horizontal line.

@@

\subsection{Method}
Data were generated from the same model and with the same type of
measurement invariance as was described in the ``Artificial Example'' section.
Sample size and magnitude
of measurement 
invariance violation were manipulated to examine power:
we examined power to detect 
invariance violations across four sample sizes ($n=50, 100,
200, 500$) and 17~magnitudes of violations.  These violations
involved the younger students' values of $\{\lambda_{11},
\lambda_{21}, \lambda_{31}\}$ deviating from the older students'
values by $d$ 
times the parameters' asymptotic standard errors (scaled by $\sqrt{n}$),
with $d = 0, 0.25, 0.5, \dots, 4$.  The 0-standard error condition was
used to study type~I error rate.

@@

\subsection{Results}
Full simulation results are presented in Figure~\ref{fig:simres}, and
the underlying numeric values for a subset of the results is additionally displayed in
Table~\ref{tab:simres}.  In describing the results, we largely refer
to the figure.

@@

To study stereotype threat in a measurement invariance framework, Study~1 of
\citeA{WicDol05} involved 295 high school students
completing three intelligence tests in two between-subjects conditions. 
Conditions were defined by whether or not students received primes about ethnic
stereotypes prior to testing. To study the data, WDH employed a
four-group, one-factor model with the three intelligence tests as manifest
variables. The groups were defined by ethnicity (majority/minority) and by the
experimental manipulation (received/did not receive stereotype prime). Results
indicated that the intelligence tests lacked measurement invariance, with the
minority group receiving stereotype primes being particularly different from the
other three groups on the most difficult intelligence test.  In the
example below, we employ one of the models used by WDH.
Our $V$ is participants' aptitudes, as measured by their grade point averages (GPAs).  The GPAs were unused in the original analyses.

\subsection{Method}
WDH tested a series of confirmatory factor models for
various types of measurement invariance.  We focus on the model used
in their Step~5b (see pp.~703--704 of their paper), which involved
across-group restrictions on the factor loadings, unique
variances, intercepts, and factor variances.  These parameters were
typically restricted to be equal across groups, though a subset of the
parameters were allowed to be group-specific upon examination of
modification indices.    The model
provided a reasonable fit to the data, as judged by examination of
$\chi^2$, RMSEA (root-mean-square error of approximation), and CFI
(comparative fit index) statistics.  While the model included four
groups as described above, we focus only on the submodel for the minority
group that received stereotype primes, which is pictured in Figure~\ref{fig:wpath}.  In the figure, paths with dashed lines and red parameters signify group-specific parameters.  These include the factor mean $\eta$, numerical factor loading $\lambda_{\text{num}}$, numerical intercept $\mu_{\text{num}}$, and numerical unique variance $\psi_{\text{num}}$.  Other parameters (not displayed) were restricted to be equal across all four groups, with the exception of the factor variance.  The factor variance was restricted to be equal within both minority groups, and separately within both majority groups.

\begin{figure}
\caption{Path diagram representing the submodel for the ``minority, stereotype'' group in model 5b of WDH.  Dashed paths and red parameters are group specific.  Other parameters (not displayed) were restricted to be equal across all four groups, with the factor variance only being restricted across the two minority groups.}
\label{fig:wpath}
\includegraphics[height=3.5in]{wicherts_msp.pdf}
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for the four group-specific parameters across GPA
  in the ``minority, stereotype prime'' group
  using data from Study 1 of Wicherts et al.\ (2005). %% \citeA{WicDol05} does not work in caption
  The solid gray horizontal lines correspond to the critical value of
  the doule-max test at $\alpha = 0.05$.}
\label{fig:cusumgpa}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-4-plot, fig=TRUE, height=8, width=5.5>>=
gefp_4_bound <- get_boundary(gefp_4_info, maxBB)
plot(gefp_4_info$process,
  ylim = c(-2.15, 2.15), main = "DM, k* = 4", xlab = "GPA",
  ylab = expression(lambda[num], eta, mu[num], psi[num]),
  panel = function(x, y, ...) {
    abline(h = 0)
    lines(gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(-gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(x, y, ...)
  })

@@

The step functions from~\eqref{eq:h1*} were employed in the current paper to
highlight connections with nested model
comparison, but the proposed tests typically have non-trivial power for
all (non-constant) deviation patterns $\theta(v_i)$ (considered in
Equation~\eqref{eq:h1}). Thus, the tests will also have power for linear
deviations from parameter constancy, although other techniques may 
perform better in this particular case.  Examples include the 
class of moderated factor models \cite{BauHus09,MolDol10,Pur02}, 
whereby continuous moderators are allowed to linearly impact model
parameters (Purcell also considers quadratic effects of the
moderators). Moreover, if the change in
parameters is continuous but not exactly linear -- e.g., a sigmoidal
shift from one set of parameters to another one -- then the simple
single shift model in~\eqref{eq:h1*} may be a useful first
approximation and may have better power than linear techniques
(depending on how clearly the two regimes are separated).

>>>>

To employ the LR~test for assessing measurement invariance, model
parameters are estimated separately for a certain number of subgroups
of the data (with some parameters potentially restricted to be equal across
subgroups).  For ease of exposition, we describe
the case where there are no such parameter restrictions; as shown in
the example and simulation below, however, it is
straightforward to extend all methods to the more general case.
After fitting the model to each subgroup, the sum of maximized likelihoods 
from the subgroups are compared with the original maximized full-sample likelihood
in a $\chi^2$~test. For the special case of two subgroups,
the alternative $H_1^*$ from (\ref{eq:h1*}) with fixed and prespecified $\nu$ is adopted
and the null hypothesis $H_0$ from (\ref{eq:h0}) reduces to ${\bm \theta}^{(A)} = {\bm \theta}^{(B)}$.
The parameter estimates $\hat {\bm \theta}^{(A)}$ can then be obtained from the
observations $i = 1, \dots, m$, say, for which $v_i \le \nu$. Analogously,
$\hat {\bm \theta}^{(B)}$ is obtained by maximizing the likelihood for the
observations $i = m + 1, \dots, n$, for which $v_i > \nu$. The LR~test statistic
for the given threshold $\nu$ is then
\begin{equation} \label{eq:lr}
  \mathit{LR}(\nu) ~=~ -2 \left[
         \ell(\hat {\bm \theta}; x_1, \dots, x_n)
   ~-~ \{\ell(\hat {\bm \theta}^{(A)}; x_1, \dots, x_m)
    +    \ell(\hat {\bm \theta}^{(B)}; x_{m+1}, \dots, x_n)\}
    \right],
\end{equation}
which, when the null hypothesis holds, has an asymptotic $\chi^2$ with degrees
of freedom equal to the number of parameters in~${\bm \theta}$.

@@

For application to measurement invariance, the most important
theoretical result involves the fact that, under $H_0$ in \eqref{eq:h0}, the
\emph{cumulative score process} converges to a specific asymptotic
process. The $k$-dimensional cumulative score process is defined as
\begin{equation} \label{eq:cumscore}
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n t \rfloor} {\bm s}(\hat {\bm \theta}; x_i)
  \qquad (0 \le t \le 1)
\end{equation}
where $\lfloor nt \rfloor$ is the integer part of $nt$ and
$\hat I$ is some consistent estimate of the covariance matrix of the scores,
e.g., their outer product or the observed information matrix.
The $k$~dimensions of the 
process arise from the fact that a separate cumulative score is
maintained for each of the $k$ model parameters.  
As the equation shows, the cumulative score process adds subsets of
casewise score contributions across individuals along the ordering
w.r.t.\ the variable $V$ of interest.  At $t=1/n$, only the
first individual's contribution enters into the summation; at
$t=2/n$, the first two individuals' contributions enter into the
summation, etc., until $t = n/n$
where all contributions enter into the summation. Thus, due to (\ref{eq:ml1}),
the cumulative score process always equals zero at $t = 0$ and
returns to zero at $t = 1$. Furthermore, multiplication by
$\hat {\bm I}^{-1/2}$ ``decorrelates'' the $k$~cumulative score
processes, such that each univariate process (i.e., each
process for a single model parameter) is unrelated to (and asymptotically independent of) all other
processes. Therefore, this cumulative process ${\bm B}(t; \hat {\bm \theta})$
accomplishes the challenges discussed at the beginning of this section:
it makes use of the ordering of the observations by taking cumulative sums,
and it decorrelates the contributions of the $k$ different parameters.

@@

The empirical cumulative score process from (\ref{eq:cumscore}) can also be viewed
as an $n \times k$ matrix with elements ${\bm B}(i/n; \hat {\bm \theta})_j$ that
we also denote ${\bm B}(\hat {\bm \theta})_{ij}$ below for brevity.
That is, assuming that individuals are ordered by $V$, the first row
of ${\bm B}(\hat {\bm \theta})$ corresponds to the first individual's
decorrelated scores.  The second row of ${\bm B}(\hat {\bm \theta})$ corresponds to
the sum of the first two individuals' scores, etc., until the last row
of ${\bm B}(\hat {\bm \theta})$ corresponds to the sum of all
individuals' scores (which will be a row of zeroes).
Under this setup, each column of ${\bm B}(\hat {\bm \theta})$
converges to a univariate Brownian bridge and pertains 
to a single factor analysis parameter. To carry out a test of $H_0$, the process/matrix
needs to be aggregated to a scalar test statistic by collapsing across
rows (individuals) and columns (parameters) of the matrix. The asymptotic distribution
of this test statistic is then easily obtained by applying the same
aggregation to the asymptotic Brownian bridge \cite{HjoKon02,ZeiHor07},
so that corresponding $p$-values can be derived.

As argued above, no single aggregation function will have high power for all
conceivable patterns of measurement invariance ${\bm \theta}(V)$, while any
(reasonable) aggregation function will have non-trivial power under $H_1$.
Thus, various aggregation strategies should be employed depending on which
pattern ${\bm \theta}(V)$ is most plausible (because the exact pattern
is typically
unknown). A particularly agnostic aggregation strategy is to reject $H_0$
if any component of the cumulative score process ${\bm B}(t; \hat {\bm \theta})$
strays ``too far'' from zero at any time, i.e., if the double maximum statistic
\begin{equation}
    \label{eq:dmax}
    \mathit{DM} = \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,
\end{equation}
becomes large.  In determining where in ${\bm B}(\hat {\bm \theta})$
this maximum occurred, we are able to determine  
threshold(s) of parameter change (over the individuals $i = 1, \dots, n$)
and the parameter(s) affected by it (over $j = 1, \dots, k$). This test is
especially useful for visualization, as the cumulative score process for
each individual parameter can be displayed along with the appropriate
critical value. For an example of this see Figure~\ref{fig:lambdas}
(in the ``Example'' section) which shows the cumulative score processes for three
factor loadings along with the critical values at 5\% level.

@@

As pointed out above, specification of the asymptotic distribution under $H_0$ for the test
statistics from the previous section is straightforward: it is simply the aggregation
of the asympotic process ${\bm B}^0(t)$ \cite{HjoKon02,ZeiHor07}. Thus, $\mathit{DM}$
from (\ref{eq:dmax}) converges to $\sup_t || {\bm B}^0(t) ||_{\infty}$, where
$||\cdot||_\infty$ denotes the maximum norm. Similarly, $\mathit{CvM}$
from (\ref{eq:cvm}) converges to $\int_0^1 || {\bm B}^0(t) ||_2^2 d t$, where $||\cdot||_2$
denotes the Euclidean norm. Finally, $\max \mathit{LM}$ from (\ref{eq:maxlm}) -- and analogously
the maximum Wald and LR~tests -- converges to $\sup_t (t (1-t))^{-1} || {\bm B}^0(t) ||_2^2$
(which can also be interpreted as the maximum of a tied-down Bessel process, as pointed out
previously).

@@

\subsubsection{Overall Results}
Test statistics for the hypotheses \eqref{eq:simh0} and~\eqref{eq:h0}
are displayed in Figure~\ref{fig:cusumex}. Each panel displays a test 
statistic's fluctuation across values of student age, with the first
column containing tests of~\eqref{eq:simh0} and the second column
containing tests of~\eqref{eq:h0}.  The solid
horizontal 
lines represent critical values for $\alpha= 0.05$, and
the Cram\'{e}r-von Mises panels 
also contain a dashed line depicting the value of the test
statistic $\mathit{CvM}$ (test statistics for the others are simply 
the maxima of the processes).
In other words, for panels in the first and third rows, \eqref{eq:simh0} is
rejected if the process crosses the horizontal line.  For panels in
the second row, \eqref{eq:h0} is rejected if the dashed horizontal line is
higher than the solid horizontal line.

@@

\subsection{Method}
Data were generated from the same model and with the same type of
measurement invariance violation as was described in the ``Artificial Example'' section.
Sample size and magnitude
of measurement 
invariance violation were manipulated to examine power:
we examined power to detect 
invariance violations across four sample sizes ($n=50, 100,
200, 500$) and 17~magnitudes of violations.  These violations
involved the younger students' values of $\{\lambda_{11},
\lambda_{21}, \lambda_{31}\}$ deviating from the older students'
values by $d$~times the parameters' asymptotic standard errors (scaled by $\sqrt{n}$),
with $d = 0, 0.25, 0.5, \dots, 4$.  The $d=0$ standard error condition was
used to study type~I error rate.

@@

\subsection{Results}
Full simulation results are presented in Figure~\ref{fig:simres}, and
the underlying numeric values for a subset of the results are additionally displayed in
Table~\ref{tab:simres}.  In describing the results, we largely refer
to the figure.

@@

To study stereotype threat in a measurement invariance framework, Study~1 of
\citeA{WicDol05} involved 295 high school students
completing three intelligence tests in two between-subjects conditions. 
Conditions were defined by whether or not students received primes about ethnic
stereotypes prior to testing. To study the data, WDH employed a
four-group, one-factor model with the three intelligence tests as manifest
variables. The groups were defined by ethnicity (majority/minority) and by the
experimental manipulation (received/did not receive stereotype prime). Results
indicated that the intelligence tests lacked measurement invariance, with the
minority group receiving stereotype primes being particularly different from the
other three groups on the most difficult intelligence test.  In the
example below, we employ one of the models used by WDH.
Our $V$ is participants' aptitudes, as measured by their grade point averages (GPAs) which were unused in the original analyses.

\subsection{Method}
WDH tested a series of confirmatory factor models for
various types of measurement invariance.  We focus on the model used
in their Step~5b (see pp.~703--704 of their paper), which involved
across-group restrictions on the factor loadings, unique
variances, intercepts, and factor variances.  These parameters were
typically restricted to be equal across groups, though a subset of the
parameters were allowed to be group-specific upon examination of
modification indices.    The model
provided a reasonable fit to the data, as judged by examination of
$\chi^2$, RMSEA (root-mean-square error of approximation), and CFI
(comparative fit index) statistics.  While the model included four
groups as described above, we focus only on the submodel for the minority
group that received stereotype primes, which is pictured in Figure~\ref{fig:wpath}.  In the figure, paths with dashed lines and \fixme{red} parameters signify group-specific parameters.  These include the factor mean $\eta$, numerical factor loading $\lambda_{\text{num}}$, numerical intercept $\mu_{\text{num}}$, and numerical unique variance $\psi_{\text{num}}$.  Other parameters (not displayed) were restricted to be equal across all four groups, with the exception of the factor variance.  The factor variance was restricted to be equal within both minority groups, and separately within both majority groups.

\begin{figure}
\caption{Path diagram representing the submodel for the ``minority, stereotype prime'' group in model 5b of WDH.  Dashed paths and red (FIXME) parameters are group specific.  Other parameters (not displayed) were restricted to be equal across all four groups, with the factor variance only being restricted across the two minority groups.}
\label{fig:wpath}
\includegraphics[height=3.5in]{wicherts_msp.pdf}
\end{figure}

@@

\begin{figure}
\centering
\caption{Cumulative score processes for the four group-specific parameters across GPA
  in the ``minority, stereotype prime'' group
  using data from Study 1 of WDH. %% \citeA{WicDol05} does not work in caption
  The solid gray horizontal lines correspond to the critical value of
  the doule-max test at $\alpha = 0.05$.}
\label{fig:cusumgpa}
\setkeys{Gin}{width=0.8\textwidth}
<<gefp-4-plot, fig=TRUE, height=8, width=5.5>>=
gefp_4_bound <- get_boundary(gefp_4_info, maxBB)
plot(gefp_4_info$process,
  ylim = c(-2.15, 2.15), main = "DM, k* = 4", xlab = "GPA",
  ylab = expression(lambda[num], eta, mu[num], psi[num]),
  panel = function(x, y, ...) {
    abline(h = 0)
    lines(gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(-gefp_4_bound, col = "slategray", lwd = 1.7)
    lines(x, y, ...)
  })

@@

The step functions from~\eqref{eq:h1*} were employed in the current paper to
highlight connections with nested model
comparison, but the proposed tests typically have non-trivial power for
all (non-constant) deviation patterns $\theta(v_i)$ (considered in
Equation~\eqref{eq:h1}). Thus, the tests will also have power for linear
deviations from parameter constancy, although other techniques may 
perform better in this particular case.  Examples include the 
class of moderated factor models \cite{BauHus09,MolDol10,Pur02}, 
whereby continuous moderators are allowed to linearly impact model
parameters (\citeNP{Pur02}, also considers quadratic effects of the
moderators). Moreover, if the change in
parameters is continuous but not exactly linear -- e.g., a sigmoidal
shift from one set of parameters to another one -- then the simple
single shift model in~\eqref{eq:h1*} may be a useful first
approximation and may have better power than linear techniques
(depending on how clearly the two regimes are separated).
